---
title: "Phylofactor Trait Analysis"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}

library(tidyverse)

load("~/Desktop/BDEL/BZDEL/Data/Meta Analyses Papers/Olival_Appended_Dataset.Rdata")

olival_et_al_2017_appended <- olival_et_al_2017_appended %>%
  filter(!is.na(phyloBin))
```

# Current State of the Dataset
### Transmission Mode 

Transmission mode is coming from the Geohegan & Holmes paper. We have `r olival_et_al_2017_appended %>% filter(IsZoonotic==1) %>%  select(Transmission_Mode, virus_name_dc) %>% unique () %>% dplyr::summarise(counts =n())` and of those `r olival_et_al_2017_appended %>% filter(IsZoonotic==1) %>% select(Transmission_Mode, virus_name_dc) %>% unique () %>% filter(is.na(Transmission_Mode)) %>% dplyr::summarise(counts =n()) ` are NA. So, for the majority of observations (`r round(100* olival_et_al_2017_appended %>% filter(IsZoonotic==1) %>% select(Transmission_Mode, virus_name_dc) %>% unique () %>% filter(is.na(Transmission_Mode)) %>% dplyr::summarise(counts =n()) /  olival_et_al_2017_appended %>% filter(IsZoonotic==1) %>% select(Transmission_Mode, virus_name_dc) %>% unique () %>% dplyr::summarise(counts =n()),1)`%) we do not have data on this transmission mode yet.

## Envelope Status 

Envelope status is coming from the original Olival paper. We have `r olival_et_al_2017_appended %>% select(vEnvelope, virus_name_dc) %>% unique () %>% dplyr::summarise(counts =n())` and of those `r olival_et_al_2017_appended %>% select(vEnvelope, virus_name_dc) %>% unique () %>% filter(is.na(vEnvelope)) %>% dplyr::summarise(counts =n()) ` are NA. So, (`r round(100* olival_et_al_2017_appended %>% select(vEnvelope, virus_name_dc) %>% unique () %>% filter(is.na(vEnvelope)) %>% dplyr::summarise(counts =n()) /  olival_et_al_2017_appended %>% select(vEnvelope, virus_name_dc) %>% unique () %>% dplyr::summarise(counts =n()),1)`%) observations are missing data on envelope status, hooray! 

```{r, echo=FALSE, message=FALSE, warning=FALSE}

y <- olival_et_al_2017_appended %>% 
  select(Transmission_Mode, virus_name_dc) %>% 
  unique () 

#y <- olival_et_al_2017_appended %>%
#  filter(is.na(Envelope_Status))
envelope_breakdown <- olival_et_al_2017_appended %>%
    select(phyloBin, vEnvelope, virus_name_dc, IsZoonotic) %>%
    unique() %>%
    unite(phylo_envelope, phyloBin, vEnvelope) 

ggplot(olival_et_al_2017_appended) +
    geom_bar(aes(x=as.factor(vEnvelope), fill= as.factor(IsZoonotic)))  +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggplot(envelope_breakdown) +
    geom_bar(aes(x=as.factor(phylo_envelope), fill= as.factor(IsZoonotic)))  +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

## Vector Status 

Transmission mode is coming from the Geohegan & Holmes paper. We have `r olival_et_al_2017_appended %>% filter(IsZoonotic==1) %>% select(vVectorYNna, virus_name_dc) %>% unique () %>% dplyr::summarise(counts =n())` and of those `r olival_et_al_2017_appended %>% filter(IsZoonotic==1) %>% select(vVectorYNna, virus_name_dc) %>% unique () %>% filter(is.na(vVectorYNna)) %>% dplyr::summarise(counts =n()) ` are NA. So, for(`r round(100* olival_et_al_2017_appended %>% filter(IsZoonotic==1) %>% select(vVectorYNna, virus_name_dc) %>% unique () %>% filter(is.na(vVectorYNna)) %>% dplyr::summarise(counts =n()) /  olival_et_al_2017_appended %>% filter(IsZoonotic==1) %>% select(vVectorYNna, virus_name_dc) %>% unique () %>% dplyr::summarise(counts =n()),1)`%) of observations we do not have data on vector transmission yet.

```{r, echo=FALSE, message=FALSE, warning=FALSE}

vector_breakdown <- olival_et_al_2017_appended %>%
    select(phyloBin, vVectorYNna, virus_name_dc, IsZoonotic) %>%
    unique() %>%
    unite(phylo_vVectorYNna, phyloBin, vVectorYNna) 

ggplot(olival_et_al_2017_appended) +
    geom_bar(aes(x=as.factor(vVectorYNna), fill= as.factor(IsZoonotic)))  +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggplot(vector_breakdown) +
    geom_bar(aes(x=as.factor(phylo_vVectorYNna), fill= as.factor(IsZoonotic)))  +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

## Human to Human Transmission
 
Human to Human transmission is also coming from the Geohegan & Holmes paper. We have `r olival_et_al_2017_appended %>% filter(IsZoonotic==1) %>% select(Humany_to_Human, virus_name_dc) %>% unique () %>% dplyr::summarise(counts =n())` and of those `r olival_et_al_2017_appended %>% filter(IsZoonotic==1) %>% select(Humany_to_Human, virus_name_dc) %>% unique () %>% filter(is.na(Humany_to_Human)) %>% dplyr::summarise(counts =n()) ` are NA. So, for(
`r round(100* olival_et_al_2017_appended %>% filter(IsZoonotic==1) %>% select(Humany_to_Human, virus_name_dc) %>% unique () %>% filter(is.na(Humany_to_Human)) %>% dplyr::summarise(counts =n()) /  olival_et_al_2017_appended %>% filter(IsZoonotic==1) %>% select(Humany_to_Human, virus_name_dc) %>% unique () %>% dplyr::summarise(counts =n()),1)`%) of observations we do not have data on vector transmission yet.

```{r, echo=FALSE, message=FALSE, warning=FALSE}

y <- olival_et_al_2017_appended %>%
  filter(Transmission_Mode=='bodily fluids') %>%
  select(virus_name_dc, hHostNameFinal, IsZoonotic)



```